Circumnuclear Media and Accretion Rates of Quiescent Supermassive Black Holes

### 2015-04-07 ms.tex
fixed Ia discussion later

Our criterion for cooling instability given by equations \eqref{eq:cooling2} and ~\eqref{eq:Mdotmax} is no longer valid if $\rs$ is regulated to $\rIa$ instead of being determined by equation \eqref{eq:stag_simple}.  However, by re-evaluating the density and temperature of gas at $\rIa$ using equations  (\ref{eq:rhors2}) and (\ref{eq:Tanalytic}) for $v_w = v_{w}^{\rm Ia}$, we the ratio of heating to radiative cooling at $\rIa$ is given by
\begin{align}
&\left.\frac{\dot{q}_{\rm heat}}{\left|\dot{q}_{\rm rad}\right|}\right|_{\rIa}\simeq
\begin{cases}
  90 \Mbheight^{1.3} (\epsilon_{\rm Ia}/0.4)^{1.7} (\tau_{\star}/\th)^{-3}
  \eta_{0.02}^{-2.7} & \text{core}\nonumber\\
   90 \Mbheight^{0.8} (\epsilon_{\rm Ia}/0.4)^{1.7} (\tau_{\star}/\th)^{-2.7}
  \eta_{0.02}^{-2.7} & \text{cusp}\nonumber\\
\end{cases}\\&\simeq
\begin{cases}
  90 \Mbheight^{1.3} (\epsilon_{\rm Ia}/0.4)^{1.7}
  (\tau_{\star}/\th)^{0.5}
 & \text{core}\\
   90 \Mbheight^{0.8} (\epsilon_{\rm Ia}/0.4)^{1.7} (\tau_{\star}/\th)^{0.9}
  & \text{cusp}\\
\end{cases}
\end{align}
Thus, only for high mass galaxies and/or relatively old stellar populations is Ia heating is sufficient to prevent cooling instabilities on scales of the stagnation radius.

@@

An important caveat, however, is that $\vw$ is only a few times higher than the stellar velocity dispersion, i.e. $\zeta \sim \zeta_c$.  This implies that the true stagnation radius could be larger than our analytic estimates, potentially resulting in thermal instability, especially for core galaxies with large break radii ($\S\ref{sec:mdot}$).

@@

For low mass black holes, blow out from SN Ia radius caps the SMBH
accretion rate at a value $\dot{M}_{\rm Ia} \sim
10^{-6}-10^{-5}M_{\odot}$ yr$^{-1}$ (green lines;
eq.~[\ref{eq:eddr_Ia}]), which is however not low enough to prevent
otherwise thermally-unstable flow at smaller radii (see discussion
following eq.~[\ref{eq:eddr_Ia}]).  On the other hand, for higher
$M_{\bullet}$ the stagnation resides close to the Ia radius so Ia
heating contributes to the steady gas heating rate, however the energy
released by a single Ia is insufficient to unbind gas from the stellar
bulge (eq.~[\ref{eq:blowout}]).

>>>>

Our criterion for cooling instability given by equations \eqref{eq:cooling2} and ~\eqref{eq:Mdotmax} is no longer valid if $\rs$ is regulated to $\rIa$ instead of being determined by equation \eqref{eq:stag_simple}.  However, by re-evaluating the density and temperature of gas at $\rIa$ using equations  (\ref{eq:rhors2}) and (\ref{eq:Tanalytic}) for $v_w = v_{w}^{\rm Ia}$, we the ratio of heating to radiative cooling at $\rIa$ is given by
\begin{align}
\label{eq:heatingratioIa}
&\left.\frac{\dot{q}_{\rm heat}}{\left|\dot{q}_{\rm rad}\right|}\right|_{\rIa}\simeq
\begin{cases}
  90 \Mbheight^{1.3} (\epsilon_{\rm Ia}/0.4)^{1.7} (\tau_{\star}/\th)^{-3}
  \eta_{0.02}^{-2.7} & \text{core}\nonumber\\
   90 \Mbheight^{0.8} (\epsilon_{\rm Ia}/0.4)^{1.7} (\tau_{\star}/\th)^{-2.7}
  \eta_{0.02}^{-2.7} & \text{cusp}\nonumber\\
\end{cases}\\&\simeq
\begin{cases}
  90 \Mbheight^{1.3} (\epsilon_{\rm Ia}/0.4)^{1.7}
  (\tau_{\star}/\th)^{0.5}
 & \text{core}\\
   90 \Mbheight^{0.8} (\epsilon_{\rm Ia}/0.4)^{1.7} (\tau_{\star}/\th)^{0.9}
  & \text{cusp}\\
\end{cases}
\end{align}
Thus, for high mass galaxies and/or relatively old stellar populations, Ia heating is capable of preventing cooling instabilities on scales of the stagnation radius $(\dot{q}_{\rm heat}/|\dot{q}_{\rm rad})_{\rIa})$. However, for lower mass galaxies or younger stellar populations, Ia blow-out is itself unlikely to prevent cooling instabilities.

@@

An important caveat, however, is that $\vw$ is only a few times higher than the stellar velocity dispersion, i.e. $\zeta \sim \zeta_c$ (eq.~[\ref{eq:zetacrit}]).  This implies that the true stagnation radius and SMBH accretion rate could be larger than our analytic estimates, potentially resulting in thermal instability, especially for core galaxies with large break radii ($\S\ref{sec:mdot}$).

@@

For low mass black holes, blow out from SN Ia radius caps the SMBH
accretion rate at a value $\dot{M}_{\rm Ia} \sim
10^{-6}-10^{-5}M_{\odot}$ yr$^{-1}$ (green lines;
eq.~[\ref{eq:eddr_Ia}]), which is however not low enough to prevent
otherwise thermally-unstable flow at smaller radii (eq.~[\ref{eq:heatingratioIa}]).  On the other hand, for higher
$M_{\bullet}$ the stagnation resides close to the Ia radius so Ia
heating contributes to the steady gas heating rate, even if the energy
released by a single Ia is insufficient to unbind gas from the stellar
bulge (eq.~[\ref{eq:blowout}]).
